Effectiveness of COVID-19 Booster doses in England: a sequential trials study in OpenSAFELY

### 2022-02-13 manuscript/draft-manuscript.Rmd

# where are the outputs (ie the inputs for this manuscript!) saved?
#output_dir_os <- here("output", "manuscript-objects")
output_dir_os <- here("released_output", "manuscript-objects")

@@

  outcomes <- 
    c(
      "postest",
      "covidemergency"#,
      #"covidadmitted",
      #"covidcc",
      #"coviddeath"
    ) %>% 
    set_names(., .)

@@

hazards <- read_csv(fs::path(output_dir_os, "effects.csv")) %>%
  mutate(
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
  )

#cmlinc3 <- cmlinc %>% filter(model==3)
#riskdiff3 <- riskdiff %>% filter(model==3)
hazards3 <- hazards %>% filter(model==3)

@@

The national COVID-19 vaccination programme in the United Kingdom began a new phase in September 2021 with the administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course (that is, "fully vaccinated" with a complete first and second dose). Initially offered to those at highest risk of severe disease, eligibility was extended to the entire adult population by the end of the year. In the vast majority of cases, BNT162b2 and mRNA-1273 vaccines were used and administered without consideration of the type of primary course.

@@

@@

## Study population and design

We studied adults (18 and over) in England who had received a complete primary COVID-19 vaccination course (two doses) of BNT162b2 or ChAdOx1-S and who were registered at a GP practice using TPP's SystmOne clinical information system. People were not eligible for inclusion if: they were a vaccinated health care worker, resident in a care or nursing home, medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); had received heterologous (i.e., mixed) first and second vaccine dose; information on sex, ethnicity, deprivation, and geographical region was not known.

After exclusions for missing values on demographic variables, there were no missing values in remaining variables as they were defined by the presence or absence of clinical codes or events in the patient record.

We used a sequential trials approach designed to emulate a sequence of two-arm parallel randomised controlled trials recruiting participants between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` inclusive. For each trial, the treatment arm comprises people who received a booster dose on that day, and the control arm comprises the same number of unboosted people selected using exact 1:1 matching without replacement. Trial arms are matched on primary course vaccine type (BNT162b2, ChAdOx1-S, or mRNA-1273), the calendar week of the second vaccine dose, geographical region, prior infection status, and vaccine priority group (as determined by the Joint Committee for Vaccine and immunisation expert working group). People selected as controls are not eligible to be included again as a control in a subsequent trial, but are eligible for selection into the treatment arm of a subsequent trial. Unmatched booster recipients were not included.

Additional eligiblity criteria based on sufficiently many people being vaccinated on each day within region, priority group, and second dose week strata were also used to ensure unboosted people were both eligible and able to receive a booster dose at the time of potential recruitment. The trial sequence was terminated early if fewer than 10% of treated people were successfully matched. See supplementary materials for further details.

@@

The following five outcomes were considered: positive SARS-CoV-2 test; COVID-19 A&E attendance; unplanned COVID-19 hospital admission; COVID-19 critical care admission; COVID-19 death. Positive SARS-CoV-2 tests were identified using SGSS records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow tests are included, without differentiation between symptomatic and asymptomatic infection. COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc. Unplanned COVID-19 hospital admissions and critical care admissions were identified using HES in-patient hospital records with U07.1 or U07.2 reason for admission ICD-10 codes. COVID-19 deaths with ICD-10 codes (as above) mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death) were included.

@@

Participant characteristics used describe the cohort and for confounder adjustment include: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintiles), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), rurality (urban conurbation, urban city or town, rural town or village), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission. All characteristics were ascertained as at the trial recruitment date (or study start date if assumed to be fixed from that date onwards).

@@

Time zero was the trial recruitment day, and trial participants were followed up for no more than 12 weeks from this day. Follow-up was censored earlier than this at `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`, death, or GP practice de-registration. Additionally for each matched pair, follow-up for both the boosted and unboosted person was censored if the control is boosted.

Data from each trial were pooled and analysed using Cox regression with period-specific hazard ratios for treatment, in two-week periods. Baseline hazards were allowed to differ for participants differing by recruitment day, region, priority group, and week and type of second vaccine dose, via Cox stratification. Other variables were included as covariates. A robust variance estimator was used to account for participant clustering.

@@

A total of `r filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment arm (supplementary Tables S1 and S2). Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1), with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms (Table 2).

@@

Characteristics at recruitment between treatment and controls arms are provided in supplementary materials (Table S2). Characteristics were (??) well-balanced between trial arms.

@@

#### Figure 1: Booster vaccination uptake, and matching success

Booster vaccination uptake over the duration of the study period, and the proportion of those that were matched for modelling.

@@

#### Table 1: Baseline characteristics

Participant characteristics as on the day of recruitment into the treatment or control arms, stratified by treatment and outcome type.

@@

table1 <- 
  table1_body %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))

@@

```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
incidences %>%
  filter(fup_period == "All") %>%
  left_join(
    matchsummary %>% select(treatment, eligible_1=total_treated, recruit_prop_1 = prop_recruited),
    by=c("treatment")
  ) %>%
  select(
    treatment_descr, 
    outcome_descr,
    eligible_1,
    n_1,
    recruit_prop_1,
    q_1,
    rate_1,
    q_0, 
    rate_0,
    rr, 
    rrCI
  ) %>%
  gt(
    groupname_col = c("treatment_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    outcome_descr = "Outcome",

@@

In the BNT162b2 and mRNA-1273 analyses respectively, matched participants contributed `r filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="postest", fup_period=="all") %>% pull(events)` positive SARS-CoV-2 tests, `r filter(incidences, treatment=="pfizer", outcome=="covidattendance", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidattendance", fup_period=="all") %>% pull(events)` COVID-19 A&E attendances, `r filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` COVID-19 hospital admissions, `r filter(incidences, treatment=="pfizer", outcome=="covidcc", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidcc", fup_period=="all") %>% pull(events)` COVID-19 ciritical care admissions, and `r filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="coviddeath", fup_period=="all") %>% pull(events)` COVID-19 deaths (Table 2).

During the fifth and sixth week after recruitment (days 28-41) the estimated hazard ratio comparing a BNT162b2 booster dose to no dose for a positive test was `r filter(hazards, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` (Figure 2). For mRNA-1273, the corresponding estimated hazard ratios were `r filter(hazards, outcome=="postest", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="moderna", term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`.

Estimated effectiveness by primary vaccine course...

@@

For each outcome based on the fully adjusted model, the period-specific hazard ratio for boosting versus not boosting is shown. Models with less extensive confounder adjustment are provided in supplementary materials (supplementary Figure SX).

@@

  plot_hazards <-
    hazards3 %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(outcome_descr), cols=vars(treatment_descr), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.01, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=postbaselinecuts, limits=c(min(postbaselinecuts), max(postbaselinecuts)+1), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_bw()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),

>>>>

# where are the outputs (ie the inputs for this manuscript!) saved?
output_dir_os <- here("output", "manuscript-objects")
#output_dir_os <- here("released_output", "manuscript-objects")

@@

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidcc",
      "coviddeath"
    ) %>% 
    set_names(., .)

@@

hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv")) %>%
  mutate(
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
  )

hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")) %>%
  mutate(
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
  )

#cmlinc3 <- cmlinc %>% filter(model==3)
#riskdiff3 <- riskdiff %>% filter(model==3)
hazards3_none <- hazards_none %>% filter(model==3)
hazards3_vax12_type <- hazards_vax12_type %>% filter(model==3)

@@

The national COVID-19 vaccination programme in the United Kingdom began a new phase in September 2021 with the administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course (that is, "fully vaccinated" with a complete first and second dose). Initially offered to those at highest risk of severe disease, eligibility was extended to the entire adult population by the end of the year.

@@

In the vast majority of cases, BNT162b2 and mRNA-1273 vaccines were used, and administered without consideration of the type of primary course. BNT162b2 was used from 16 September 2022, and mRNA-1273 from X onwards. Consequently, available follow-up of BNT162b2 recipients is longer, and so in this early analysis we restrict to effectiveness of this vaccine. An analysis of mRNA-1273 booster effectiveness is planned when sufficient follow-up has accrued.

@@

## Study design and population

We used a sequential trials approach designed to emulate a sequence of daily two-arm parallel randomised controlled trials. For each trial, the treatment arm comprised people who received a BNT162b2 booster dose on that day, and the control arm comprised the same number of randomly selected unboosted people, matched on key characteristics described below.

Trial recruitment was considered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` inclusive. Trial eligibility on each day of recruitment was as follows: adults (18 and over) who were registered at a GP practice using TPP's SystmOne clinical information system; received a primary COVID-19 vaccination course of either two BNT162b2 doses or two ChAdOx1-S doses (mRNA-1273 was not considered due to small numbers); not a vaccinated health care worker, not a resident in a care or nursing home, and not medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); no evidence of SARS-CoV-2 infection or COVID-19 disease within 90 days prior to recruitment; not undergoing an unplanned hospital admission; information on sex, ethnicity, deprivation, and geographical region was complete.

Additional eligiblity criteria based on sufficiently many people being vaccinated on each day within region, priority group, and second dose week strata were also used to ensure unboosted people were both eligible and able to receive a booster dose at the time of potential recruitment. The trial sequence was terminated early if nobody was eligible for the treatment arm or fewer than 10% of eligible treated people were successfully matched. See supplementary materials for further details.

The control arms were selected using exact 1:1 matching without replacement. Participants were matched on primary course vaccine type (BNT162b2 or ChAdOx1-S), the calendar week of the second vaccine dose, geographical region, prior infection, immunosuppression, and vaccine priority group (as determined by the Joint Committee for Vaccine and immunisation expert working group). People selected as controls were not eligible to be included again as a control in a subsequent trial, but were eligible for selection into the treatment arm of a subsequent trial. Any unmatched booster recipients were dropped.

After exclusions for missing values on demographic variables, there were no missing values in remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.

@@

The following three outcomes were considered: positive SARS-CoV-2 test; unplanned COVID-19 hospital admission; COVID-19 death. Positive SARS-CoV-2 tests were identified using SGSS records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow tests are included, without differentiation between symptomatic and asymptomatic infection. 
Unplanned COVID-19 hospital admissions were identified using HES in-patient hospital records with U07.1 or U07.2 reason for admission ICD-10 codes. COVID-19 deaths with the same COVID-19 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death) were included.

<!-- COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc. -->

COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases.

@@

Participant characteristics used describe trial participants and for confounder adjustment include: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintiles), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), rurality (urban conurbation, urban city or town, rural town or village), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission. All characteristics were ascertained as at the trial recruitment date (or study start date if assumed to be fixed from that date onwards).

@@

Time zero was the trial recruitment day, and trial participants were followed up for no more than 10 weeks from this day. Follow-up was censored earlier than this at `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`, death, or GP practice de-registration. Additionally for each matched pair, follow-up for both the boosted and unboosted person was censored if the control is boosted.

Trials were pooled and analysed using Cox regression with period-specific hazard ratios for treatment, in two-week periods. Baseline hazards were allowed to differ for participants differing by recruitment day, region, and week and type of second vaccine dose, via Cox stratification. Other variables were included as covariates. A robust variance estimator was used to account for participant clustering.

@@

A total of `r filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment arm (supplementary Tables S1 and S2). Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1).

@@

Participant characteristics at trial recruitment between trial arms were often well-balanced (Table 1), with some notable exceptions. Unboosted controls were typically... These unbalanced characteristics were adjusted for in the Cox regression models.

There were `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms (Table 2).

@@

#### Figure 1: Booster vaccination uptake, treatment arm eligiblity and matching success

Booster vaccination uptake over the duration of the study period, and the number of boosted people who were eligible and matched.

@@

#### Table 1: Participant characteristics

Participant characteristics as on the day of recruitment into the treatment or control arm.

@@

table1 <- 
  table1_body %>%
  #filter(outcome %in% outcomes) %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))

@@

```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
incidences %>%
  filter(
    fup_period == "All",
    outcome %in% c("postest", "covidadmitted", "coviddeath")
  ) %>%
  left_join(
    matchsummary %>% select(treatment, eligible_1=total_treated, recruit_prop_1 = prop_recruited),
    by=c("treatment")
  ) %>%
  select(
    treatment_descr, 
    outcome_descr,
    eligible_1,
    n_1,
    recruit_prop_1,
    q_1,
    rate_1,
    q_0, 
    rate_0,
    rr, 
    rrCI
  ) %>%
  gt(
    groupname_col = c("treatment_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    outcome_descr = "Outcome",

@@

Trial participants contributed `r filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="all") %>% pull(events)` positive SARS-CoV-2 tests, `r filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` COVID-19 hospital admissions, and `r filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="all") %>% pull(events)` COVID-19 deaths (Table 2).

<!-- `r filter(incidences, treatment=="pfizer", outcome=="covidattendance", fup_period=="all") %>% pull(events)` COVID-19 A&E attendances, `r filter(incidences, treatment=="pfizer", outcome=="covidcc", fup_period=="all") %>% pull(events)` COVID-19 ciritical care admissions, -->

During the fifth and sixth week after recruitment (days 28-41) the estimated hazard ratio comparing a BNT162b2 booster dose to no dose was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).

<!-- `r filter(hazards_none, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 A&E attendances, `r filter(hazards_none, outcome=="covidcc", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 critical care admission,  -->

<!-- For mRNA-1273, the corresponding estimated hazard ratios were `r filter(hazards, outcome=="postest", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="moderna", term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`.  -->

For the same period, the estimated effectiveness for BNT162b2 and ChAdOx1-S primary courses respectively were  `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).

<!--  `r filter(hazards_vax12_type, outcome=="covidemergency", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="emergency", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 A&E attendances, -->
<!-- `r filter(hazards_vax12_type, outcome=="covidcc", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidcc", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 critical care admission, -->

@@

For each outcome based on the fully adjusted model, the period-specific hazard ratio for boosting with BNT162b2 versus not boosting is shown, stratified by primary course. Models with less extensive confounder adjustment are provided in supplementary materials (supplementary Figure SX).

@@

  plot_hazards <-
    hazards3_vax12_type %>%
    filter(outcome %in% c("postest", "covidadmitted", "coviddeath")) %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_descr), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.01, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=postbaselinecuts, limits=c(min(postbaselinecuts), max(postbaselinecuts)+1), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_bw()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),
